---
title: "Regular Expressions in R"
author: "Daloha Rodr&iacute;guez-Molina"
date: "1/19/2018"
output: html_document
---

# Regular Expressions in R
## Course: Intermediate R

### Functions to use
?grep     
?grepl     
?sub     
?gsub     

### 1. grep
Identifies a regular expression and returns a numeric vector indicating the position of the element that matches the regex.

**Usage**: `grep(pattern = <regex>, x = <string>)`
```{r}
animals <- c("cat", "moose", "impala", "ant", "kiwi")
grep(pattern = "a", x = animals)
[1] 1 3 4
```

### 2. grepl
Same as `grep` but returns a logical vector.

**Usage**: `grepl(pattern = <regex>, x = <string>)`
```{r}
animals <- c("cat", "moose", "impala", "ant", "kiwi")
grepl(pattern = "a", x = animals)
[1]  TRUE FALSE  TRUE  TRUE FALSE
```

Doing `grep(pattern = "a", x = animals)` is the same as doing `which(grepl(pattern = "a", x = animals))`:
```{r}
grep(pattern = "a", x = animals)
[1] 1 3 4

which(grepl(pattern = "a", x = animals))
[1] 1 3 4
```
#### Options for `grep`, `grepl` and regex in general:
* match regex to the beginning of an element: Use `^regex`
```{r}
grepl(pattern = "^a", x = animals)
[1] FALSE FALSE FALSE  TRUE FALSE
```
* match regex to the end of an element: Use `regex$`
```{r}
grepl(pattern = "a$", x = animals)
[1] FALSE FALSE  TRUE FALSE FALSE
```

### 3. sub
Replaces matches with other strings.     
**Usage**: `sub(pattern = <regex>, replacement = <str>, x = <str>)`
```{r}
sub(pattern = "a", replacement = "o", x = animals)
[1] "cot"    "moose"  "impola" "ont"    "kiwi" 
```
**Warning**: `sub()` only looks for the first match in the string! Therefore, "impala" was changed to "impola". The last "a" is conserved.     
**Solution**: `gsub()`

### 4. gsub
Does the same as `sub()` but replaces all instances in a string.
```{r}
gsub(pattern = "a", replacement = "o", x = animals)
[1] "cot"    "moose"  "impolo" "ont"    "kiwi"  
```
#### Options for `sub`, `gsub` and regex in general:
* The OR metacharacter: Takes into account two regex to replace:
```{r}
gsub(pattern = "a|i", replacement = "_", x = animals)
[1] "c_t"    "moose"  "_mp_l_" "_nt"    "k_w_" 

gsub(pattern = "a|i|o", replacement = "_", x = animals)
[1] "c_t"    "m__se"  "_mp_l_" "_nt"    "k_w_" 
```

### Working with emails (copied from the course exercises)
You can use the caret, `^`, and the dollar sign, `$` to match the content located in the start and end of a string, respectively. This could take us one step closer to a correct pattern for matching only the ".edu" email addresses from our list of emails. But there's more that can be added to make the pattern more robust:

`@`, because a valid email must contain an at-sign.
`.*`, which matches any character (.) zero or more times (*). Both the dot and the asterisk are metacharacters. You can use them to match any character between the at-sign and the ".edu" portion of an email address.
`\\.edu$`, to match the ".edu" part of the email at the end of the string. The `\\` part escapes the dot: it tells R that you want to use the `.` as an actual character.

#### Exercise
```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for .edu addresses more robustly
grepl("@.*\\.edu$", emails)
[1]  TRUE FALSE FALSE FALSE  TRUE FALSE

# Use grep() to match for .edu addresses more robustly, save result to hits
hits <- grep("@.*\\.edu$", emails)

# Subset emails using hits
emails[hits]
[1] "john.doe@ivyleague.edu"   "quant@bigdatacollege.edu"
```

#### Exercise with sub
```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "global@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use sub() to convert the email domains to datacamp.edu
sub("@.*\\.edu$", "@datacamp.edu", emails)
```

#### Moar info (copied from DataCamp):
Regular expressions are a typical concept that you'll learn by doing and by seeing other examples. Before you rack your brains over the regular expression in this exercise, have a look at the new things that will be used:

`.*`: A usual suspect! It can be read as "any character that is matched zero or more times".
`\\s`: Match a space. The "s" is normally a character, escaping it (`\\`) makes it a metacharacter.
`[0-9]+`: Match the numbers 0 to 9, at least once (+).
`([0-9]+)`: The parentheses are used to make parts of the matching string available to define the replacement. The `\\1` in the replacement argument of `sub()` gets set to the string that is captured by the regular expression `[0-9]+`.

```{r}
awards <- c("Won 1 Oscar.",
  "Won 1 Oscar. Another 9 wins & 24 nominations.",
  "1 win and 2 nominations.",
  "2 wins & 3 nominations.",
  "Nominated for 2 Golden Globes. 1 more win & 2 nominations.",
  "4 wins & 1 nomination.")

sub(".*\\s([0-9]+)\\snomination.*$", "\\1", awards)
[1] "Won 1 Oscar." "24"           "2"            "3"            "2"           
[6] "1"  
```

Great! Can you explain why all of this happened? The `([0-9]+)` selects the entire number that comes before the word “nomination” in the string, and the entire match gets replaced by this number because of the `\\1` that reference to the content inside the parentheses. The next video will get you up to speed with times and dates in R!











