---
title: "Regular Expressions in R"
author: "Daloha Rodr&iacute;guez-Molina"
date: "1/19/2018"
output: html_document
---

# Regular Expressions in R
## Course: Intermediate R

### Functions to use
?grep     
?grepl     
?sub     
?gsub     

### 1. grep
Identifies a regular expression and returns a numeric vector indicating the position of the element that matches the regex.

**Usage**: `grep(pattern = <regex>, x = <string>)`
```{r}
animals <- c("cat", "moose", "impala", "ant", "kiwi")
grep(pattern = "a", x = animals)
[1] 1 3 4
```

### 2. grepl
Same as `grep` but returns a logical vector.

**Usage**: `grepl(pattern = <regex>, x = <string>)`
```{r}
animals <- c("cat", "moose", "impala", "ant", "kiwi")
grepl(pattern = "a", x = animals)
[1]  TRUE FALSE  TRUE  TRUE FALSE
```

Doing `grep(pattern = "a", x = animals)` is the same as doing `which(grepl(pattern = "a", x = animals))`:
```{r}
grep(pattern = "a", x = animals)
[1] 1 3 4

which(grepl(pattern = "a", x = animals))
[1] 1 3 4
```
#### Options for `grep`, `grepl` and regex in general:
* match regex to the beginning of an element: Use `^regex`
```{r}
grepl(pattern = "^a", x = animals)
[1] FALSE FALSE FALSE  TRUE FALSE
```
* match regex to the end of an element: Use `regex$`
```{r}
grepl(pattern = "a$", x = animals)
[1] FALSE FALSE  TRUE FALSE FALSE
```

### 3. sub
Replaces matches with other strings.     
**Usage**: `sub(pattern = <regex>, replacement = <str>, x = <str>)`
```{r}
sub(pattern = "a", replacement = "o", x = animals)
[1] "cot"    "moose"  "impola" "ont"    "kiwi" 
```
**Warning**: `sub()` only looks for the first match in the string! Therefore, "impala" was changed to "impola". The last "a" is conserved.     
**Solution**: `gsub()`

### 4. gsub
Does the same as `sub()` but replaces all instances in a string.
```{r}
gsub(pattern = "a", replacement = "o", x = animals)
[1] "cot"    "moose"  "impolo" "ont"    "kiwi"  
```
#### Options for `sub`, `gsub` and regex in general:
* The OR operator: Takes into account two regex to replace:
```{r}
gsub(pattern = "a|i", replacement = "_", x = animals)
[1] "c_t"    "moose"  "_mp_l_" "_nt"    "k_w_" 

gsub(pattern = "a|i|o", replacement = "_", x = animals)
[1] "c_t"    "m__se"  "_mp_l_" "_nt"    "k_w_" 
```












