# Shiny

## Reactivity
**Goal**: to create a checkbox that shows a table if checked. We need:
* ui: add an input widget (the checkbox)
* ui: add an output defining where the table should appear.
* server: add a reactive expression that creates the table *if* the checkbox is checked.

### Tip for table outputs
`DT::dataTableOutput(outputId = "id")`

### Sample code for the server
``` {r}
# print data table if checked
output$id <- DT::renderDataTable ({
  if(input$show_data) {
    DT:datatable(data = movies %>% select(1:7),
    options = list(pageLenght = 10),
    rownames = FALSE)
  }
})
```

### Tip: it is possible to use static html statements
`HTML(paste0("The dataset has ", nrow(movies), "observations."))`

## The `req` function
If an error such as `Error: size is not a numeric or integer vector` appears, it's probably because Shiny is expecting a numerical input and this has been deleted.
To solve it, use the `req` function. we need to hold back the output from being calculated if the input is missing. The req function is the simplest and best way to do this, it ensures that values are available ("truthy") before proceeding with a calculation or action. If any of the given values is not truthy, the operation is stopped by raising a "silent" exception (not logged by Shiny, nor displayed in the Shiny app's UI).

**Before:**
```{r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    movies_sample <- movies %>%
      sample_n(input$n) %>%
      select(title:studio)
    DT::datatable(data = movies_sample, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

**After:**
```{r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    req(input$n)
    movies_sample <- movies %>%
      sample_n(input$n) %>%
      select(title:studio)
    DT::datatable(data = movies_sample, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

## The `selectize` function
* The selectize option suggests names of studios as you type them.
* The server function needs some changes to go with `multiple = TRUE` in the `selectInput` function:

**Before**:
```{r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    movies_from_selected_studios <- movies %>%
      filter(studio == input$studio) %>%
      select(title:studio)
    DT::datatable(data = movies_from_selected_studios, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

* Add a call to the `req` function in the server, just like you did in the previous exercise but this time requiring that `input$studio` be available. Update the call to `filter()` to use the logical operator `%in%` (instead of `==`) so that it works when multiple studios are selected.

**After:**
``` {r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    req(input$studio)
    movies_from_selected_studios <- movies %>%
      filter(studio %in% input$studio) %>%
      select(title:studio)
    DT::datatable(data = movies_from_selected_studios, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

## Render functions
The expression in the render function must be wrapped in curly brackets.

## Tables from selected area in plots
We'll **brush** some observations in the plot and generate a reactive table that shows information about the selected data points.
1. ui: add functionality to `plotOutput` to select plots via brushing.
```{r}
# Show scatterplot with brushing capability
plotOutput(outputId = "scatterplot", brush = "plot_brush")
```

2. ui: add output defining where the data table should appear.
```{r}
# Show data table
DT::dataTableOutput(outputId = "moviestable")
```
3. server: add a reactive expression that creates the data table from the selected points.
```{r}
# print data table
output$moviestable <- DT::renderDataTable({
  brushedPoints(movies, input$plot_brush) %>%
    select(title, audience_score, critics_score)
})
```

We can also **hover** over data points.
1. ui: add functionality to `plotOutput` to select plots via hovering.
```{r}
plotOutput(outputId = "scatterplot", hover = "plot_hover"),
```
2. ui: add output defining where the data table should appear.
```{r}      
dataTableOutput(outputId = "moviestable")
```
3. server: add a reactive expression that creates the data table from the hovered points.
```{r}
# Create data table
  output$moviestable <- DT::renderDataTable({
    nearPoints(movies, input$plot_hover, xvar = input$x, yvar = input$y) %>% 
      select(title, audience_score, critics_score)
  })
```

## Building an app
The goal in this exercise is to develop an app where the user selects two variables and their relationship is visualized with a scatterplot, and averages of both variables are reported as well as the output of the linear regression predicting the variable on the y-axis from the variable in the x-axis. The code on the right only does some of these things.

```{r}
library(shiny)
library(dplyr)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# UI
ui <- fluidPage(
  sidebarLayout(
    
    # Input(s)
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y",
                  label = "Y-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"),
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x",
                  label = "X-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"),
                  selected = "critics_score")
      
    ),
    
    # Output(s)
    mainPanel(
      plotOutput(outputId = "scatterplot"),
      textOutput(outputId = "avg_x"), # avg of x
      textOutput(outputId = "avg_y"), # avg of y
      verbatimTextOutput(outputId = "lmoutput") # regression output
    
  )
))

# Server
server <- function(input, output) {
  
  # Create scatterplot
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y)) +
      geom_point()
  })
  
  # Calculate average of x
  output$avg_x <- renderText({
    avg_x <- movies %>% pull(input$x) %>% mean() %>% round(2)
    paste("Average", input$x, "=", avg_x)
  })
  
  # Calculate average of y
  output$avg_y <- renderText({
    avg_y <- movies %>% pull(input$y) %>% mean() %>% round(2)
    paste("Average", input$y, "=", avg_y)
  })
  
  # Create regression output
  output$lmoutput <- renderPrint({
    x <- movies %>% pull(input$x)
    y <- movies %>% pull(input$y)
    summ <- summary(lm(y ~ x, data = movies)) 
    print(summ, digits = 3, signif.stars = FALSE)
  })
  
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

## Downloading data
```{r}
library(shiny)
library(dplyr)
library(readr)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# UI
ui <- fluidPage(
  sidebarLayout(
    
    # Input(s)
    sidebarPanel(
      
      # Select filetype
      radioButtons(inputId = "filetype",
                   label = "Select filetype:",
                   choices = c("csv", "tsv"),
                   selected = "csv"),
      
      # Select variables to download
      checkboxGroupInput(inputId = "selected_var",
                  label = "Select variables:",
                  choices = names(movies),
                  selected = c("title"))
      
    ),
    
    # Output(s)
    mainPanel(
      HTML("Select filetype and variables, then hit 'Download data'."),
      downloadButton("download_data", "Download data")
    )
  )
)

# Server
server <- function(input, output) {
  
  # Download file
  output$download_data <- downloadHandler(
    filename = function() {
      paste0("movies.", input$filetype)
      },
    content = function(file) { 
      if(input$filetype == "csv"){ 
        write_csv(movies %>% select(input$selected_var), file) 
        }
      if(input$filetype == "tsv"){ 
        write_tsv(movies %>% select(input$selected_var), file) 
        }
    }
  )
  
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```
