# Shiny

## Reactivity
**Goal**: to create a checkbox that shows a table if checked. We need:
* ui: add an input widget (the checkbox)
* ui: add an output defining where the table should appear.
* server: add a reactive expression that creates the table *if* the checkbox is checked.

### Tip for table outputs
`DT::dataTableOutput(outputId = "id")`

### Sample code for the server
``` {r}
# print data table if checked
output$id <- DT::renderDataTable ({
  if(input$show_data) {
    DT:datatable(data = movies %>% select(1:7),
    options = list(pageLenght = 10),
    rownames = FALSE)
  }
})
```

### Tip: it is possible to use static html statements
`HTML(paste0("The dataset has ", nrow(movies), "observations."))`

## The `req` function
If an error such as `Error: size is not a numeric or integer vector` appears, it's probably because Shiny is expecting a numerical input and this has been deleted.
To solve it, use the `req` function. we need to hold back the output from being calculated if the input is missing. The req function is the simplest and best way to do this, it ensures that values are available ("truthy") before proceeding with a calculation or action. If any of the given values is not truthy, the operation is stopped by raising a "silent" exception (not logged by Shiny, nor displayed in the Shiny app's UI).

**Before:**
```{r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    movies_sample <- movies %>%
      sample_n(input$n) %>%
      select(title:studio)
    DT::datatable(data = movies_sample, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

**After:**
```{r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    req(input$n)
    movies_sample <- movies %>%
      sample_n(input$n) %>%
      select(title:studio)
    DT::datatable(data = movies_sample, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

## The `selectize` function
* The selectize option suggests names of studios as you type them.
* The server function needs some changes to go with `multiple = TRUE` in the `selectInput` function:

**Before**:
```{r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    movies_from_selected_studios <- movies %>%
      filter(studio == input$studio) %>%
      select(title:studio)
    DT::datatable(data = movies_from_selected_studios, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

* Add a call to the `req` function in the server, just like you did in the previous exercise but this time requiring that `input$studio` be available. Update the call to `filter()` to use the logical operator `%in%` (instead of `==`) so that it works when multiple studios are selected.

**After:**
``` {r}
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    req(input$studio)
    movies_from_selected_studios <- movies %>%
      filter(studio %in% input$studio) %>%
      select(title:studio)
    DT::datatable(data = movies_from_selected_studios, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}
```

## Render functions
The expression in the render function must be wrapped in curly brackets.

## Tables from selected area in plots
We'll **brush** some observations in the plot and generate a reactive table that shows information about the selected data points.
1. ui: add functionality to `plotOutput` to select plots via brushing.
```{r}
# Show scatterplot with brushing capability
plotOutput(outputId = "scatterplot", brush = "plot_brush")
```

2. ui: add output defining where the data table should appear.
```{r}
# Show data table
DT::dataTableOutput(outputId = "moviestable")
```
3. server: add a reactive expression that creates the data table from the selected points.
```{r}
# print data table
output$moviestable <- DT::renderDataTable({
  brushedPoints(movies, input$plot_brush) %>%
    select(title, audience_score, critics_score)
})
```

We can also **hover** over data points.
1. ui: add functionality to `plotOutput` to select plots via hovering.
```{r}
plotOutput(outputId = "scatterplot", hover = "plot_hover"),
```
2. ui: add output defining where the data table should appear.
```{r}      
dataTableOutput(outputId = "moviestable")
```
3. server: add a reactive expression that creates the data table from the hovered points.
```{r}
# Create data table
  output$moviestable <- DT::renderDataTable({
    nearPoints(movies, input$plot_hover, xvar = input$x, yvar = input$y) %>% 
      select(title, audience_score, critics_score)
  })
```

## Building an app
The goal in this exercise is to develop an app where the user selects two variables and their relationship is visualized with a scatterplot, and averages of both variables are reported as well as the output of the linear regression predicting the variable on the y-axis from the variable in the x-axis. The code on the right only does some of these things.

```{r}
library(shiny)
library(dplyr)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# UI
ui <- fluidPage(
  sidebarLayout(
    
    # Input(s)
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y",
                  label = "Y-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"),
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x",
                  label = "X-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"),
                  selected = "critics_score")
      
    ),
    
    # Output(s)
    mainPanel(
      plotOutput(outputId = "scatterplot"),
      textOutput(outputId = "avg_x"), # avg of x
      textOutput(outputId = "avg_y"), # avg of y
      verbatimTextOutput(outputId = "lmoutput") # regression output
    
  )
))

# Server
server <- function(input, output) {
  
  # Create scatterplot
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y)) +
      geom_point()
  })
  
  # Calculate average of x
  output$avg_x <- renderText({
    avg_x <- movies %>% pull(input$x) %>% mean() %>% round(2)
    paste("Average", input$x, "=", avg_x)
  })
  
  # Calculate average of y
  output$avg_y <- renderText({
    avg_y <- movies %>% pull(input$y) %>% mean() %>% round(2)
    paste("Average", input$y, "=", avg_y)
  })
  
  # Create regression output
  output$lmoutput <- renderPrint({
    x <- movies %>% pull(input$x)
    y <- movies %>% pull(input$y)
    summ <- summary(lm(y ~ x, data = movies)) 
    print(summ, digits = 3, signif.stars = FALSE)
  })
  
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

## Downloading data
```{r}
library(shiny)
library(dplyr)
library(readr)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# UI
ui <- fluidPage(
  sidebarLayout(
    
    # Input(s)
    sidebarPanel(
      
      # Select filetype
      radioButtons(inputId = "filetype",
                   label = "Select filetype:",
                   choices = c("csv", "tsv"),
                   selected = "csv"),
      
      # Select variables to download
      checkboxGroupInput(inputId = "selected_var",
                  label = "Select variables:",
                  choices = names(movies),
                  selected = c("title"))
      
    ),
    
    # Output(s)
    mainPanel(
      HTML("Select filetype and variables, then hit 'Download data'."),
      downloadButton("download_data", "Download data")
    )
  )
)

# Server
server <- function(input, output) {
  
  # Download file
  output$download_data <- downloadHandler(
    filename = function() {
      paste0("movies.", input$filetype)
      },
    content = function(file) { 
      if(input$filetype == "csv"){ 
        write_csv(movies %>% select(input$selected_var), file) 
        }
      if(input$filetype == "tsv"){ 
        write_tsv(movies %>% select(input$selected_var), file) 
        }
    }
  )
  
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

## More reactivity

### Reactlog
It's a graphical representation of the dependency structure of reactive expressions in Shiny. It also gives very detailed information about what is happening under the hood as Shiny evaluates an app. To view:
* In a fresh R session, run `options(shiny.reactlog = TRUE)`
* Then, launch your application as you normally would
* In the app, press CMD + F3

## Stop, trigger, delay
* `observe()`: observers are useful for their side effects, but they don't calculate anything. 
* `reactive()`: reactive expressions return values, and have no side effects.    
** A side effect is e.g. download a file.**
* `observeEvent()`: Used to trigger a reaction. To **perform an action** in response to an event. E.g. Download a data frame when user presses action button:
```{r}
# ui
actionButton(inputId = "write_csv", label = "Write CSV")

# server
observeEvent(input$write_csv, {
              filename <- paste0("movies_",
                          str_replace_all(Sys.time(), ":|\", "_"),
                          ".csv")
              write_csv(movies_sample(), path = filename)
              }                
```
* `eventReactive()`: Used to delay a reaction. Used to **create a calculated value** that only updates in response to an event. E.g. update calculations for a plot based on sample size inputted by the user only when the user clicks on action button "Calculate!". **Note**: The first argument in `eventReactive()` is the expression of the event to condition on, the second argument is the side-effect-producing action that you want the app to take when the first expression is invalidated, and this argument goes in curly braces.

`observe()` and `reactive()` automatically trigger whatever they access.
`observeEvent()` and `eventReactive()` need to be explicitely told what triggers them.

* `isolate()`: to stop and **isolate reactions**. E.g. Update a plot and its title only when inputs other than `input$plot_title` changes:
```{r}
output$scatterplot <- renderPlot({
  ggplot(data = movies_subset(), aes_string(x = input$x, y = input$y)) +
    geom_point() +
    labs(title = isolate({ input$plot_title )} )
})
```
## Trigger with `observeEvent()`
In this app we want two things to happen when an action button is clicked: (1) A message printed to the console stating how many records are shown and (2) A table output of those records.

## Reactivity recap
### Three lessons:
1. Reactives are equivalent to no-argument functions. Think about them as functions and as variables that can depend on user inputs or other reactives. They're lazily evaluated: They will only evaluate when their inputs change, not each time they're called.
2. Reactives are for reactive values and expressions, observers are for their side effects.
3. Do not define a reactive inside a render function. 

## Layout with html
* Tags: `names(tags)`
* Conditional Panels. The first argument in this function is the `condition`, which is a JavaScript expression that will be evaluated repeatedly to determine whether the panel should be displayed. The `condition` should be stated as `"input.show_data == true"`. Note that we use `true` instead of `TRUE` because this is a Javascript expression, as opposed to an R expression.
```{r}
library(shiny)
library(ggplot2)
library(stringr)
library(dplyr)
library(DT)
library(tools)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  titlePanel("Movie browser, 1970 - 2014", windowTitle = "Movies"),
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs
    sidebarPanel(
    
      wellPanel(
        h3("Plotting"),      # Third level header: Plotting
        
        # Select variable for y-axis 
        selectInput(inputId = "y", 
                    label = "Y-axis:",
                    choices = c("IMDB rating" = "imdb_rating", 
                                "IMDB number of votes" = "imdb_num_votes", 
                                "Critics Score" = "critics_score", 
                                "Audience Score" = "audience_score", 
                                "Runtime" = "runtime"), 
                    selected = "audience_score"),
        
        # Select variable for x-axis 
        selectInput(inputId = "x", 
                    label = "X-axis:",
                    choices = c("IMDB rating" = "imdb_rating", 
                                "IMDB number of votes" = "imdb_num_votes", 
                                "Critics Score" = "critics_score", 
                                "Audience Score" = "audience_score", 
                                "Runtime" = "runtime"), 
                    selected = "critics_score"),
        
        # Select variable for color
        selectInput(inputId = "z", 
                    label = "Color by:",
                    choices = c("Title Type" = "title_type", 
                                "Genre" = "genre", 
                                "MPAA Rating" = "mpaa_rating", 
                                "Critics Rating" = "critics_rating", 
                                "Audience Rating" = "audience_rating"),
                    selected = "mpaa_rating"),
        
        hr(),
        
        # Set alpha level
        sliderInput(inputId = "alpha", 
                    label = "Alpha:", 
                    min = 0, max = 1, 
                    value = 0.5),
        
        # Set point size
        sliderInput(inputId = "size", 
                    label = "Size:", 
                    min = 0, max = 5, 
                    value = 2),
        
        # Enter text for plot title
        textInput(inputId = "plot_title", 
                  label = "Plot title", 
                  placeholder = "Enter text to be used as plot title")
        
      ),
      
      wellPanel(
        # Header
        h3("Subsetting and sampling"),
        
        # Select which types of movies to plot
        checkboxGroupInput(inputId = "selected_type",
                           label = "Select movie type(s):",
                           choices = c("Documentary", "Feature Film", "TV Movie"),
                           selected = "Feature Film"),
        
        # Select sample size
        numericInput(inputId = "n_samp", 
                     label = "Sample size:", 
                     min = 1, max = nrow(movies), 
                     value = 50)        
      ),
      
      wellPanel(
        # Show data table
        checkboxInput(inputId = "show_data",
                      label = "Show data table",
                      value = TRUE)
      ),
      
      # Built with Shiny by RStudio
      br(),
      h5("Built with",
         img(src = "https://www.rstudio.com/wp-content/uploads/2014/04/shiny.png", height = "30px"),
         "by",
         img(src = "https://www.rstudio.com/wp-content/uploads/2014/07/RStudio-Logo-Blue-Gray.png", height = "30px"),
         "."),
      
      width = 5
      
    ),
    
    # Output:
    mainPanel(
      
      # Show scatterplot
      h3("Scatterplot"),    # Horizontal line for visual separation
      plotOutput(outputId = "scatterplot"),
      br(),                 # Single line break for a little bit of visual separation
      
      # Print number of obs plotted
      h4(uiOutput(outputId = "n")),
      br(), br(),           # Two line breaks for a little bit of visual separation
      
      # Show data table
      conditionalPanel("input.show_data == true", h3("Data table")),     # Third level header: Data table
      DT::dataTableOutput(outputId = "moviestable"),
      
      width = 7
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create a subset of data filtering for selected title types
  movies_subset <- reactive({
    req(input$selected_type) # ensure availablity of value before proceeding
    filter(movies, title_type %in% input$selected_type)
  })
  
  # Update the maximum allowed n_samp for selected type movies
  observe({
    updateNumericInput(session, 
                       inputId = "n_samp",
                       value = min(50, nrow(movies_subset())),
                       max = nrow(movies_subset())
    )
  })
  
  # Create new df that is n_samp obs from selected type movies
  movies_sample <- reactive({ 
    req(input$n_samp) # ensure availablity of value before proceeding
    sample_n(movies_subset(), input$n_samp)
  })
  
  # Create scatterplot object the plotOutput function is expecting 
  output$scatterplot <- renderPlot({
    ggplot(data = movies_sample(), aes_string(x = input$x, y = input$y,
                                              color = input$z)) +
      geom_point(alpha = input$alpha, size = input$size) +
      labs(x = toTitleCase(str_replace_all(input$x, "_", " ")),
           y = toTitleCase(str_replace_all(input$y, "_", " ")),
           color = toTitleCase(str_replace_all(input$z, "_", " ")),
           title = toTitleCase(input$plot_title))
  })
  
  # Print number of movies plotted 
  output$n <- renderUI({
    types <- movies_sample()$title_type %>% 
      factor(levels = input$selected_type) 
    counts <- table(types)
    
    HTML(paste("There are", counts, input$selected_type, "movies in this dataset. <br>"))
  })
  
  # Print data table if checked
  output$moviestable <- DT::renderDataTable(
    if(input$show_data){
      DT::datatable(data = movies_sample()[, 1:7], 
                    options = list(pageLength = 10), 
                    rownames = FALSE)
    }
  )
  
  }

# Create Shiny app object
shinyApp(ui = ui, server = server)
```

## Conditional Tabs
Remember earlier we used a conditional panel to display/hide the heading "Data table" depending on whether the "Show data table" checkbox input was selected? We might want to do something similar with the "Data" tab now. And we can achieve this using conditional tabs, via the `showTab()` and `hideTab()` functions.
```{r}
    # ui
    mainPanel(
      
      tabsetPanel(id = "tabspanel",
                  type = "tabs",
                  tabPanel(title = "Plot", 
                           plotOutput(outputId = "scatterplot"),
                           br(),
                           h4(uiOutput(outputId = "n"))),
                  tabPanel(title = "Data", 
                           br(),
                           DT::dataTableOutput(outputId = "moviestable")),
                  # New tab panel for Codebook
                  tabPanel("Codebook", 
                           br(),
                           DT::dataTableOutput("codebook"))
                  
      )
    )
    
    # server
      # Display data table tab only if show_data is checked
  observeEvent(input$show_data, {
               if(input$show_data){
                 showTab(inputId = "tabspanel", target = "Data", select = TRUE)
               } else {
                 hideTab(inputId = "tabspanel", target = "Data")
               }
  })
```

## Whole app with theme
```{r}
library(shiny)
library(shinythemes)
library(readr)
library(ggplot2)
library(stringr)
library(dplyr)
library(DT)
library(tools)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))
movies_codebook <- read_csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies_codebook.csv")

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  theme = shinytheme("flatly"),
  
  titlePanel("Movie browser, 1970 - 2014", windowTitle = "Movies"),
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs
    sidebarPanel(
      
      h3("Plotting"),      # Third level header: Plotting
      
      # Select variable for y-axis 
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("IMDB rating" = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics Score" = "critics_score", 
                              "Audience Score" = "audience_score", 
                              "Runtime" = "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis 
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("IMDB rating" = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics Score" = "critics_score", 
                              "Audience Score" = "audience_score", 
                              "Runtime" = "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(inputId = "z", 
                  label = "Color by:",
                  choices = c("Title Type" = "title_type", 
                              "Genre" = "genre", 
                              "MPAA Rating" = "mpaa_rating", 
                              "Critics Rating" = "critics_rating", 
                              "Audience Rating" = "audience_rating"),
                  selected = "mpaa_rating"),
      
      hr(),
      
      # Set alpha level
      sliderInput(inputId = "alpha", 
                  label = "Alpha:", 
                  min = 0, max = 1, 
                  value = 0.5),
      
      # Set point size
      sliderInput(inputId = "size", 
                  label = "Size:", 
                  min = 0, max = 5, 
                  value = 2),
      
      # Enter text for plot title
      textInput(inputId = "plot_title", 
                label = "Plot title", 
                placeholder = "Enter text to be used as plot title"),
      
      hr(),
      
      # Header
      h3("Subsetting and sampling"),
      
      # Select which types of movies to plot
      checkboxGroupInput(inputId = "selected_type",
                         label = "Select movie type(s):",
                         choices = c("Documentary", "Feature Film", "TV Movie"),
                         selected = "Feature Film"),
      
      # Select sample size
      numericInput(inputId = "n_samp", 
                   label = "Sample size:", 
                   min = 1, max = nrow(movies), 
                   value = 50),
      
      hr(),
      
      # Show data table
      checkboxInput(inputId = "show_data",
                    label = "Show data table",
                    value = TRUE),
      
      br(),
      
      # Built with Shiny by RStudio
      br(),
      h5("Built with",
         img(src = "https://www.rstudio.com/wp-content/uploads/2014/04/shiny.png", height = "30px"),
         "by",
         img(src = "https://www.rstudio.com/wp-content/uploads/2014/07/RStudio-Logo-Blue-Gray.png", height = "30px"),
         ".")
      
    ),
    
    # Output:
    mainPanel(
      
      tabsetPanel(id = "tabspanel", type = "tabs",
                  tabPanel(title = "Plot", 
                           plotOutput(outputId = "scatterplot"),
                           br(),
                           h4(uiOutput(outputId = "n"))),
                  tabPanel(title = "Data", 
                           br(),
                           DT::dataTableOutput(outputId = "moviestable")),
                  # New tab panel for Codebook
                  tabPanel("Codebook", 
                           br(),
                           DT::dataTableOutput("codebook"))
                  
      )
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create a subset of data filtering for selected title types
  movies_subset <- reactive({
    req(input$selected_type) # ensure availablity of value before proceeding
    filter(movies, title_type %in% input$selected_type)
  })
  
  # Update the maximum allowed n_samp for selected type movies
  observe({
    updateNumericInput(session, 
                       inputId = "n_samp",
                       value = min(50, nrow(movies_subset())),
                       max = nrow(movies_subset())
    )
  })
  
  # Create new df that is n_samp obs from selected type movies
  movies_sample <- reactive({ 
    req(input$n_samp) # ensure availablity of value before proceeding
    sample_n(movies_subset(), input$n_samp)
  })
  
  # Create scatterplot object the plotOutput function is expecting 
  output$scatterplot <- renderPlot({
    ggplot(data = movies_sample(), aes_string(x = input$x, y = input$y,
                                              color = input$z)) +
      geom_point(alpha = input$alpha, size = input$size) +
      labs(x = toTitleCase(str_replace_all(input$x, "_", " ")),
           y = toTitleCase(str_replace_all(input$y, "_", " ")),
           color = toTitleCase(str_replace_all(input$z, "_", " ")),
           title = toTitleCase(input$plot_title))
  })
  
  # Print number of movies plotted 
  output$n <- renderUI({
    types <- movies_sample()$title_type %>% 
      factor(levels = input$selected_type) 
    counts <- table(types)
    
    HTML(paste("There are", counts, input$selected_type, "movies in this dataset. <br>"))
  })
  
  # Update code below to render data table regardless of current state of input$show_data
  output$moviestable <- DT::renderDataTable({
    DT::datatable(data = movies_sample()[, 1:7], 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
  # Display data table tab only if show_data is checked
  observeEvent(input$show_data, {
               if(input$show_data){
                 showTab(inputId = "tabspanel", target = "Data", select = TRUE)
               } else {
                 hideTab(inputId = "tabspanel", target = "Data")
               }
  })
  
  # Render data table for codebook
  output$codebook <- DT::renderDataTable({
    DT::datatable(data = movies_codebook,
                  options = list(pageLength = 10, lengthMenu = c(10, 25, 40)), 
                  rownames = FALSE)
  })
  
}

# Create Shiny app object
shinyApp(ui = ui, server = server)
```
