# Shiny course with Dean Attali

## Resources
* html builder functions: [link](https://www.rdocumentation.org/packages/shiny/versions/1.0.5/topics/builder)
* `colourpicker`: [link](https://github.com/daattali/colourpicker)
```{r}
library(colourpicker)

# ui
colourInput("col", "Select a colour", value = "orange")
```
* `plotly`: used for interactive plots. It has a function `ggplotly()` that converts any plot in `ggplot2` to `plotly`. [Link](https://plot.ly/r/). **WARNING**: a `plotly` plot is not a regular plot, so you'll have to define the `ui` and `server` functions like this:
```{r}
# ui
plotlyOutput("plot")

# server
renderPlotly(ggplotly(p))
```

With this example, the user will be able to zoom into regions, or see data points when hovering over the plot:
```{r}
p <- ggplot(gapminder,
            aes(gdpPercap, lifeExp)) +
            geom_point() +
            scale_x_log10()
library(plotly)
ggplotly(p)
```
* `reactive()`: [Link](https://www.rdocumentation.org/packages/shiny/versions/1.0.5/topics/reactive)


## app adding a line of best fit with a checkbox
```{r}
# Define UI for the application
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      textInput("title", "Title", "GDP vs life exp"),
      numericInput("size", "Point size", 1, 1),
      # Add a checkbox for line of best fit
      checkboxInput("fit", "Add line of best fit", value = FALSE)
    ),
    mainPanel(
      plotOutput("plot")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  output$plot <- renderPlot({
    p <- ggplot(gapminder, aes(gdpPercap, lifeExp)) +
      geom_point(size = input$size) +
      scale_x_log10() +
      ggtitle(input$title)
    
    # When the "fit" checkbox is checked, add a line
    # of best fit
    if (input$fit) {
      p <- p + geom_smooth(method = "lm")
    }
    p
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```
## app with ggplotly
```{r}
# Load the plotly package
library(plotly)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      textInput("title", "Title", "GDP vs life exp"),
      numericInput("size", "Point size", 1, 1),
      checkboxInput("fit", "Add line of best fit", FALSE),
      colourInput("colour", "Point colour", value = "blue"),
      selectInput("continents", "Continents",
                  choices = levels(gapminder$continent),
                  multiple = TRUE,
                  selected = "Europe"),
      sliderInput("years", "Years",
                  min(gapminder$year), max(gapminder$year),
                  value = c(1977, 2002))
    ),
    mainPanel(
      # Replace the `plotOutput()` with the plotly version
      plotlyOutput("plot")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  # Replace the `renderPlot()` with the plotly version
  output$plot <- renderPlotly({
    # Convert the existing ggplot2 to a plotly plot
    ggplotly({
      data <- subset(gapminder,
                     continent %in% input$continents &
                       year >= input$years[1] & year <= input$years[2])
      
      p <- ggplot(data, aes(gdpPercap, lifeExp)) +
        geom_point(size = input$size, col = input$colour) +
        scale_x_log10() +
        ggtitle(input$title)
      
      if (input$fit) {
        p <- p + geom_smooth(method = "lm")
      }
      p
    })
  })
}

shinyApp(ui = ui, server = server)
```
## app to explore data (filtering content with slider and select inputs)
```{r}
ui <- fluidPage(
  h1("Gapminder"),
  sliderInput(inputId = "life", label = "Life expectancy",
              min = 0, max = 120,
              value = c(30, 50)),
  # Add an "All" value to the continent list
  selectInput("continent", "Continent",
              choices = c("All", levels(gapminder$continent))),
  tableOutput("table")
)

server <- function(input, output) {
  output$table <- renderTable({
    data <- gapminder
    data <- subset(
      data,
      lifeExp >= input$life[1] & lifeExp <= input$life[2]
    )
    # Don't subset the data if "All" continent are chosen
    if (input$continent != "All") {
      data <- subset(
        data,
        continent == input$continent
      )
    }
    data
  })
}

shinyApp(ui, server)
```

## app with table and plot showing the same subset 
```{r}
ui <- fluidPage(
  h1("Gapminder"),
  sliderInput(inputId = "life", label = "Life expectancy",
              min = 0, max = 120,
              value = c(30, 50)),
  selectInput("continent", "Continent",
              choices = c("All", levels(gapminder$continent))),
  # Add a plot output
  plotOutput("plot"),
  tableOutput("table")
)

server <- function(input, output) {
  output$table <- renderTable({
    data <- gapminder
    data <- subset(
      data,
      lifeExp >= input$life[1] & lifeExp <= input$life[2]
    )
    if (input$continent != "All") {
      data <- subset(
        data,
        continent == input$continent
      )
    }
    data
  })

  # Create the plot render function  
  output$plot <- renderPlot({
    # Use the same filtered data that the table uses
        data <- gapminder
    data <- subset(
      data,
      lifeExp >= input$life[1] & lifeExp <= input$life[2]
    )
    if (input$continent != "All") {
      data <- subset(
        data,
        continent == input$continent
      )
    }
    ggplot(data, aes(gdpPercap, lifeExp)) +
      geom_point() +
      scale_x_log10()
  })
}

shinyApp(ui, server)
```

## adding css
It's possible to include CSS in your app by writing it in a separate file and importing it with `includeCSS()`, but in this course we will use the simpler approach of placing the CSS code inside `tags$style()` in the UI.
```{r}
my_css <- "
#download_data {
  /* Change the background colour of the download button
     to orange. */
  background: orange;

  /* Change the text size to 20 pixels. */
  font-size: 20px;
}

#table {
  /* Change the text colour of the table to red. */
  color: red;
}
"

ui <- fluidPage(
  h1("Gapminder"),
  # Add the CSS that we wrote to the Shiny app
  tags$style(my_css),
  tabsetPanel(
    tabPanel(
      title = "Inputs",
      sliderInput(inputId = "life", label = "Life expectancy",
                  min = 0, max = 120,
                  value = c(30, 50)),
      selectInput("continent", "Continent",
                  choices = c("All", levels(gapminder$continent))),
      downloadButton("download_data")
    ),
    tabPanel(
      title = "Plot",
      plotOutput("plot")
    ),
    tabPanel(
      title = "Table",
      DT::dataTableOutput("table")
    )
  )
)

server <- function(input, output) {
  filtered_data <- reactive({
    data <- gapminder
    data <- subset(
      data,
      lifeExp >= input$life[1] & lifeExp <= input$life[2]
    )
    if (input$continent != "All") {
      data <- subset(
        data,
        continent == input$continent
      )
    }
    data
  })
  
  output$table <- DT::renderDataTable({
    data <- filtered_data()
    data
  })

  output$download_data <- downloadHandler(
    filename = "gapminder_data.csv",
    content = function(file) {
      data <- filtered_data()
      write.csv(data, file, row.names = FALSE)
    }
  )

  output$plot <- renderPlot({
    data <- filtered_data()
    ggplot(data, aes(gdpPercap, lifeExp)) +
      geom_point() +
      scale_x_log10()
  })
}

shinyApp(ui, server)
```

## app with conditional panels for inputs
```{r}
ui <- fluidPage(
  h1("Word Cloud"),
  sidebarLayout(
    sidebarPanel(
      radioButtons(
        inputId = "source",
        label = "Word source",
        choices = c(
          "Art of War" = "book",
          "Use your own words" = "own",
          "Upload a file" = "file"
        )
      ),
      conditionalPanel(
        condition = "input.source == 'own'",
        textAreaInput("text", "Enter text", rows = 7)
      ),
      # Wrap the file input in a conditional panel
      conditionalPanel(
        # The condition should be that the user selects
        # "file" from the radio buttons
        condition = "input.source == 'file'",
        fileInput("file", "Select a file")
      ),
      numericInput("num", "Maximum number of words",
                   value = 100, min = 5),
      colourInput("col", "Background colour", value = "white")
    ),
    mainPanel(
      wordcloud2Output("cloud")
    )
  )
)

server <- function(input, output) {
  data_source <- reactive({
    if (input$source == "book") {
      data <- artofwar
    } else if (input$source == "own") {
      data <- input$text
    } else if (input$source == "file") {
      data <- input_file()
    }
    return(data)
  })
  
  input_file <- reactive({
    if (is.null(input$file)) {
      return("")
    }
    readLines(input$file$datapath)
  })
  
  output$cloud <- renderWordcloud2({
    create_wordcloud(data_source(), num_words = input$num,
                        background = input$col)
  })
}

shinyApp(ui = ui, server = server)
```

## Final app with isolate and action buttons
```{r}
ui <- fluidPage(
  h1("Word Cloud"),
  sidebarLayout(
    sidebarPanel(
      radioButtons(
        inputId = "source",
        label = "Word source",
        choices = c(
          "Art of War" = "book",
          "Use your own words" = "own",
          "Upload a file" = "file"
        )
      ),
      conditionalPanel(
        condition = "input.source == 'own'",
        textAreaInput("text", "Enter text", rows = 7)
      ),
      conditionalPanel(
        condition = "input.source == 'file'",
        fileInput("file", "Select a file")
      ),
      numericInput("num", "Maximum number of words",
                   value = 100, min = 5),
      colourInput("col", "Background colour", value = "white"),
      # Add a "draw" button to the app
      actionButton(inputId = "draw", label = "Draw!")
    ),
    mainPanel(
      wordcloud2Output("cloud")
    )
  )
)

server <- function(input, output) {
  data_source <- reactive({
    if (input$source == "book") {
      data <- artofwar
    } else if (input$source == "own") {
      data <- input$text
    } else if (input$source == "file") {
      data <- input_file()
    }
    return(data)
  })
  
  input_file <- reactive({
    if (is.null(input$file)) {
      return("")
    }
    readLines(input$file$datapath)
  })
  
  output$cloud <- renderWordcloud2({
    # Add the draw button as a dependency to
    # cause the word cloud to re-render on click
    input$draw
    isolate({
      create_wordcloud(data_source(), num_words = input$num,
                       background = input$col)
    })
  })
}

shinyApp(ui = ui, server = server)
```
