# Working with lists
## Course: Intermediate R - exercises

### Extract log information (2)
In the previous exercise, you wrote a function `extract_info()`, which is available on the right. To make this function really powerful, you'll want to add an additional argument, `property`, so that you can select any property (ie. `success`, `details`, or `timestamp`) from the log entries.

Next, you can use this argument to subset the list accordingly. You cannot use the `$` notation if the element you want to select is a variable and not the actual name of a list:
```
log$property # won't work
log[[property]] # will work
```
#### Exercise
* Add an additional argument to the function, called `property`.
* Adapt the function body such that `property` is used to select the correct information from each log.
* Call `extract_info()` on logs and set the property argument to `"timestamp"`.
* Call `extract_info()` on logs and set the property argument to `"success"`.

```{r}
# logs is available in your workspace

# Adapt the extract_info() function.
extract_info <- function(x, property) {
  info <- c()
  for (log in x) {
   info <- c(info, log[[property]])
  }
  return(info)
}

# Call extract_info() on logs, set property to "timestamp"
extract_info(logs, property = "timestamp")

# Call extract_info() on logs, set property to "success"
extract_info(logs, property = "success")
```

### Extract log information (4)
In the `for` loop exercises, you wrote code that extracts information on log entries that indicate a failure. This is something your `extract_info()` function can not yet do. You can already guess what the purpose of this exercise is, right?

#### Exercise
* Add an argument to your function `extract_info()`: call it `include_all`, and make it `TRUE` by default. That is, the default is to extract all log entries, whether it indicates a failure or a success.
* Change the body of your function: inside the `for` loop, add an `if` test: if `include_all` or if `!log$success`, you want to add the `log[[property]]` to the `info` vector. In all other cases, you're not adding anything to `info`. Use the `||` operator in your condition.
* Call your new `extract_info()` function on `logs`, first without any additional arguments. The default value for `include_all`, which is `TRUE`, will be used.
* Call `extract_info()` on `logs` again; this time set `include_all` to `FALSE`.

```{r}
# logs is available in your workspace

# Adapt extract_info():
# - add argument with default value
# - change function body
extract_info <- function(x, property = "success", include_all = TRUE) {
  info <- c()
  for (log in x) {
    if (include_all == TRUE || !log$success) {
      # add if construct around the line below
      info <- c(info, log[[property]])
    } 
  }
  return(info)
}

# Call extract_info() on logs, no additional arguments
extract_info(logs)

# Call extract_info() on logs, set include_all to FALSE
extract_info(logs, include_all = FALSE)
```

### Extract log information (5)
Have another look at `logs`, that is still available in your workspace. Have you noticed that the `details` element of each log entry differs between logs indicating success and failure? For successes, it's a list with a `message` element. For failures, it's a list with two elements: `message` and `location`. We've printed out `str(logs)` for you below. See the differences in structures between a successful log and an unsuccessful log:

```
 $ :List of 3
  ..$ success  : logi TRUE
  ..$ details  :List of 1
  .. ..$ message: chr "all good"
  ..$ timestamp: POSIXct[1:1], format: "2015-09-18 13:45:27"
  ```
  ```
   $ :List of 3
  ..$ success  : logi FALSE
  ..$ details  :List of 2
  .. ..$ message : chr "human error"
  .. ..$ location: chr "waste"
  ..$ timestamp: POSIXct[1:1], format: "2015-09-17 23:37:18"
  ```
At first sight, our function only allows the selection of log entry information on the first level, such as `success` and `details`. To get information that's deeper inside the log entries, such as `message`, we'll need another function, right? Nope, your function will work just fine. To select elements from embedded lists, you can use chained selection. The following code chunk uses chained selection to return the value 2:
```{r}
x <- list(a = 1, b = list(r = 2, s = 3))
x[[c("b", "r")]]
```

#### Exercise
* Use `extract_info()` to build a vector containing the `message` elements of all log entries, irrespective of whether they indicate a failure or not.
* Use `extract_info()` to build a vector containing the `location` information for log entries indicating a failure. This means you have to set `include_all = FALSE` now!

```{r}
# logs is available in your workspace

# Defition of the extract_info() function
extract_info <- function(x, property = "success", include_all = TRUE) {
  info <- c()
  for (log in x) {
    if (include_all || !log$success) {
     info <- c(info, log[[property]])
    }
  }
  return(info)
}

# Generate vector of messages
extract_info(logs, property = c("details","message"))

# Generate vector of locations for failed log entries
extract_info(logs, property = c("details","location"), include_all = FALSE)
```

### Last exercise on lists and loops: Over to you
Now that you've played around with building up a function, making it more powerful step-by-step, you're ready to write your own function from the bottom with minimal help.

As usual, `logs`, a list of lists, is available in your workspace to test your function.

#### Exercise
* Write a function, `compute_fail_pct()`, that takes a list of log entries that is formatted as the `logs` list and returns the percentage, i.e. a value between 0 and 100, of log entries that indicate a failure. If, for example, your list of logs has a length 50, and contains 5 failures, `compute_fail_pct()` should return 5/50∗100=105/50∗100=10
* Call `compute_fail_pct()` on `logs`.

```{r}
# logs is available in your workspace

# Write the function compute_fail_pct
compute_fail_pct <- function(x) {
  failure_n <- 0
  for (i in 1:length(x)) {
    if (x[[i]][[1]] == FALSE) {
    failure_n <- failure_n + 1
    }
  }
  percentage <- failure_n / length(x) * 100
  return(percentage)
}


# Call compute_fail_pct on logs
compute_fail_pct(logs)
```
